---
title: "ISRS Exercises6.1,6.2,6.4"
author: "Roymil Terrero"
date: "6/28/2019"
output: html_document
---
#coord_flip() to flip the coordinates

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

```

```{r}
b_weight <- read.table('babyweights.txt', header = TRUE)



```

#6.1 Baby weights, Part I. 

(a) Write the equation of the regression line. 
> y = 123.05 -8.94*X 

(b) Interpret the slope in this context, and calculate the predicted birth weight of babies born to smoker and non-smoker mothers. 

> The slope indicates the relation between smoker or non smoker. In this context the average weight of the baby decreases when the mother is a smoker.

> birth_weight = 123.05 - 8.94 * smoke

(c) Is there a statistically signiﬁcant relationship between the average birth weight and smoking?

> The p-values are so small that I can reject the hypotesis that there is not relationship.

#6.2 Baby weights, Part II.

(a) Write the equation of the regression line. 
> y = 120.07 - 1.93*X

(b) Interpret the slope in this context, and calculate the predicted birth weight of ﬁrst borns and others.
> In this context the slope is not significant, since, the standar error is +-1.19. We are so uncertain about the slope, that it could be zero. Additionally, the p-value tells us that perhaps we can not reject the hypotesis that not relationship.

> birth_weight = 120.07 - 1.93 * parity_i

(c) Is there a statistically signiﬁcant relationship between the average birth weight and parity?

#6.3 Baby weights, Part III. 

(a) Write the equation of the regression line that includes all of the variables.

> y = 0.44 * x1 - 3.33 * x2 - 0.01* x3 + 1.15* x4 + 0.05* x5 - 8.40* x6

(b) Interpret the slopes of gestation and age in this context.

>

(c) The coefficient for parity is different than in the linear model shown in Exercise 6.2. Why might there be a difference? 

(d) Calculate the residual for the ﬁrst observation in the data set.

>R = 120 - 200.99 = 80.99

(e) The variance of the residuals is 249.28, and the variance of the birth weights of all babies in the data set is 332.57. Calculate the R2 and the adjusted R2. 
*Note that there are 1,236 observations in the data set.*

```{r, echo=FALSE}

```

